---
title: "Regression Results"
author: "Max O'Krepki"
date: "2/10/2021"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(sf)
library(plm)
library(lmtest)
library(tidyverse)
library(knitr)
library(tmap)
library(stargazer)
library(kableExtra)
library(tmap)
library(ggplot2)
```



```{r, eval = FALSE, echo = FALSE}
df <- data.frame(x1 = 0, y1 = 0, x2 = 1, y2 = 1)

ggplot() + geom_point() + xlim(0, 5) + ylim(0, 5) + geom_segment(aes(x = x1, y = y1,xend = x2,yend = y2), data = df) + geom_label(aes(2,2))
```

Example figure
```{r}
df <- data.frame(x1 = 0, y1 = 0, x2 = 1, y2 = 1)
# View(df)
ggplot(df) + geom_segment(aes(x1,y1,x2,y2), data = df) + xlim(0, 10) + ylim(0, 10)
```


Crime figures
```{r}
# load("crime.sf.v2")
load("ba_bgs_lean_v2")
# load("ba_bgs_long_v2")
load("regression_models")
load("crime_rate_table")
load("housing_sf")
load("crime_2017_maps")
```

Have to create these here otherwise the markdown won't know where to find them. 
```{r, echo=FALSE}
property <- c("VEHICLE THEFT", "LARCENY/THEFT", "BURGLARY", "VANDALISM", "STOLEN PROPERTY", "TRESPASS", "ARSON")
violent  <- c("ASSAULT", "ROBBERY", "SEX OFFENSES, FORCIBLE")
drug     <- "DRUG/NARCOTIC"
nuisance <- c("SUSPICIOUS OCC", "DRUNKENNESS", "DISORDERLY CONDUCT", "LOITERING", "LIQUOR LAWS", "GAMBLING", "PROSTITUTION")

```


How to create this? I could do a couple of group_by operations, just send in a new filtered data frame to each. I almost want to do some sort of cbind, or even a series of left joins. I think I'm trying to prematurely optimize, just write out the long way to do it then see how I can condense, if possible. I tell you what, do the widening before kable. 
```{r, cache=TRUE, eval=FALSE}
class(crime.sf)

# Creates crime counts for all categories
crime_summary <- as.data.frame(crime.sf) %>% group_by(year) %>% summarize("crime_counts_total" = n())

# Property
crime_summary <- left_join(crime_summary, {as.data.frame(crime.sf) %>% filter(Category %in% property) %>% group_by(year) %>% summarize("crime_counts_property" = n())})

# Violent
crime_summary <- left_join(crime_summary, {as.data.frame(crime.sf) %>% filter(Category %in% violent) %>% group_by(year) %>% summarize("crime_counts_violent" = n())})

# Nuisance
crime_summary <- left_join(crime_summary, {as.data.frame(crime.sf) %>% filter(Category %in% nuisance) %>% group_by(year) %>% summarize("crime_counts_nuisance" = n())})


# Calculates population by year
pop_sum <- ba_bgs_long %>% group_by(year) %>% summarize("pop_total" = sum(population))


# Generates the crime summary table. 
crime_summary <- crime_summary %>% mutate("crime_rate_total" = crime_counts_total/(pop_sum$pop_total/100), "crime_rate_property" = crime_counts_property/(pop_sum$pop_total/100), "crime_rate_violent" = crime_counts_violent/(pop_sum$pop_total/100), "crime_rate_nuisance" = crime_counts_nuisance/(pop_sum$pop_total/100))

crime_summary <- crime_summary[,c(1,6:9)]
```

# Table - crime rates per 100 residents per year. 
```{r, cache=TRUE, eval=FALSE}
# tk - just end up saving this data frame and load from memory, it'll be easier, honestly. 
# crime_summary$year <- as.character(crime_summary$year)

crime_rate_table <- t(crime_summary)
colnames(crime_rate_table) <- crime_rate_table[1,]

crime_rate_table <- crime_rate_table[-1,]
rownames(crime_rate_table) <- c("Total Crimes", "Property", "Violent", "Nuisance")

# crime_rate_table <- as.numeric(crime_rate_table)

save(crime_rate_table, file = "crime_rate_table")
```

```{r, echo=FALSE}

knitr::kable(crime_rate_table, digits = 1, booktabs = TRUE, caption = "San Francisco reported crime rates per 100 residents by year and type") %>% kable_styling(latex_options="scale_down") %>% add_footnote("Source: San Francisco Police Department via SF Open Data Portal", notation = "none")
```


# Figure 1
Good preview of basemaps. 
https://leaflet-extras.github.io/leaflet-providers/preview/
leaflet::providers

Problematic block group:
1500000US060750176011


Right now it's all crime. At some point I'll do just property and violent?
```{r, cache=TRUE, echo=FALSE, eval=FALSE}
ba_bgs_long$year <- as.numeric(ba_bgs_long$year)

# Come up with crime rate for 2017. 
crime_2017 <- left_join(filter(ba_bgs_long, year == 2017),{ filter(as.data.frame(crime.sf), year == 2017) %>% group_by(geoid_long) %>% summarize("crime_counts" = n()) })

# 2017 Property
crime_2017_property <- left_join(filter(ba_bgs_long, year == 2017),{ filter(as.data.frame(crime.sf), year == 2017 & Category %in% property) %>% group_by(geoid_long) %>% summarize("crime_counts" = n()) })

# 2017 Violent
crime_2017_violent <- left_join(filter(ba_bgs_long, year == 2017),{ filter(as.data.frame(crime.sf), year == 2017 & Category %in% violent) %>% group_by(geoid_long) %>% summarize("crime_counts" = n()) })

# 2017 Nuisance
crime_2017_nuisance <- left_join(filter(ba_bgs_long, year == 2017),{ filter(as.data.frame(crime.sf), year == 2017 & Category %in% nuisance) %>% group_by(geoid_long) %>% summarize("crime_counts" = n()) })

# All
crime_2017 <- crime_2017 %>% mutate("crime_rate" = crime_counts/(population/100))
range(crime_2017$crime_rate)

# Property
crime_2017_property <- crime_2017_property %>% mutate("crime_rate" = crime_counts/(population/100))

# Violent
crime_2017_violent <- crime_2017_violent %>% mutate("crime_rate" = crime_counts/(population/100))

# Nuisance
crime_2017_nuisance <- crime_2017_nuisance %>% mutate("crime_rate" = crime_counts/(population/100))


crime_2017_map          <- left_join(ba_bgs_lean, crime_2017)
crime_2017_property_map <- left_join(ba_bgs_lean, crime_2017_property)
crime_2017_violent_map  <- left_join(ba_bgs_lean, crime_2017_violent)
crime_2017_nuisance_map <- left_join(ba_bgs_lean, crime_2017_nuisance)

# save(crime_2017_map, crime_2017_property_map, crime_2017_violent_map, crime_2017_nuisance_map, file = "crime_2017_maps")
```


tmaptools::palette_explorer

```{r, fig.cap="Total crimes per 100 residents, San Francisco, 2017."}
tm_shape(crime_2017_map, bbox=tmaptools::bb(matrix(c(-122.52, 37.70692295634627, -122.34, 37.8151870985090),2,2))) + tm_polygons("crime_rate", title = "", breaks = c(0,2,5,10,20,40,100,3000)) + tm_layout(bg.color = "skyblue")
```

Property
```{r, fig.cap="Property crimes per 100 residents, San Francisco, 2017."}
tm_shape(crime_2017_property_map, bbox=tmaptools::bb(matrix(c(-122.52, 37.70692295634627, -122.34, 37.8151870985090),2,2))) + tm_polygons("crime_rate", title = "", breaks = c(0,2,5,10,20,40,100,3000)) + tm_layout(bg.color = "skyblue")
```

Violent
```{r, fig.cap="Violent crimes per 100 residents, San Francisco, 2017."}
tm_shape(crime_2017_violent_map, bbox=tmaptools::bb(matrix(c(-122.52, 37.70692295634627, -122.34, 37.8151870985090),2,2))) + tm_polygons("crime_rate", title = "", breaks = c(0,2,5,10,20,40,100,3000)) + tm_layout(bg.color = "skyblue")
```

Nuisance
```{r, fig.cap="Nuisance crimes per 100 residents, San Francisco, 2017."}
tm_shape(crime_2017_nuisance_map, bbox=tmaptools::bb(matrix(c(-122.52, 37.70692295634627, -122.34, 37.8151870985090),2,2))) + tm_polygons("crime_rate", title = "", breaks = c(0,2,5,10,20,40,100,3000)) + tm_layout(bg.color = "skyblue")
```




Bottom left
37.70692295634627, -122.51057492517226

Top right
37.81518709850902, -122.36594404123402


The housing sites are all the "b"'s in housing_cleaned_lean. It's just the block groups that get sorted after that point depending on proximity to other housing sites. 

With the other stuff I'm saving, probably just easier in the end to save this as an R binary object to just load at the very end. The final draft will likely be made in a clean Rmarkdown with all of the figure and table data saved as R binary objects. Housing_cleaned_lean is just a df, I'll have to turn that into an sf then go from there. 
```{r}
analysis_sites <- housing_sf
names(analysis_sites)[3] <- "Facility Category"
```


```{r, fig.cap="Distribution of LIHTC sites in San Francisco"}
tm_shape(ba_bgs_lean, bbox=tmaptools::bb(matrix(c(-122.52, 37.70692295634627, -122.33, 37.8151870985090),2,2))) + tm_polygons(col = "MAP_COLORS", palette = "Pastel1", minimize = TRUE, alpha = 0.6) +
  tm_shape(analysis_sites) + tm_symbols(col= "Facility Category", palette = c("forestgreen","cornflowerblue"), size = 0.3)
```




Just a simple markdown to communicate intermediate results. 
# Regression table
, type = "text"
 "footnotesize", okay, looks like the small size is okay. 
```{r results='asis'}
stargazer(model_all,model_property, model_violent, model_nuisance , omit = "geoid_long", column.sep.width = "0.5pt", no.space = TRUE, font.size = "small", omit.stat = c("rsq", "n", "ser"), dep.var.labels.include = FALSE, dep.var.caption = "", column.labels = c("Total Crimes", "Property Crimes", "Violent Crimes", "Nuisance Crimes"), title = " Table tk. Regression coefficients of neighborhood crime impact variables, by crime type", header = FALSE)
```

